Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-06-30 sattose/paper.tex
T2R clones

\begin{itemize}
  \item \textbf{Considering types:} Type 2 clones do not consider types. However, this can make a code fragment very hard to refactor, as different types can describe different functional concepts. Because of this, we propose that type 2R clones should consider types like type 1R clones do.
  \item \textbf{Having a distinction between different variables:} For type 2 clones, no identifiers would be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \texttt{myMethod(var1, var2)}, or call this method like this: \texttt{myMethod(var1, var1)}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} For type 2 clones no literals would be taken into account. We agree, as when merging the clone (for example by extracting a method), we can turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design. This threshold is displayed in equation \ref{eq:type2r}.
  \item \textbf{Consider method call signatures and define a threshold for variability in method calls:} As type-2 clones allow changes in identifiers, also the names of called methods may vary. However, because of this, completely different methods can be called in cloned fragments as a result. This poses serious challenges on refactoring and makes it more disputable whether such a clone is harmful for the maintainability of the code. This is because different method identifiers can describe a completely different functionality. Therefore, we suggest considering the call signatures of cloned methods when they are compared. We can allow variability in the rest of method identifiers by passing the function as a parameter. To limit the amount of parameters required we also recommend defining a threshold for variability in method call expressions, so only a limited number of method calls can vary.
\end{itemize}

\begin{equation}\label{eq:type2r}
\text{T2R}=\frac{Diff(l) + Diff(v) + Diff(m)}{Total(t)}*100
\end{equation}
$$Diff = \text{Number that differ from other clone instances}$$
$$Total = \text{Total number in the clone instance}$$
$$l = \text{Literals in clone}$$
$$v = \text{Variables in clone}$$
$$m = \text{Method calls in clone}$$
$$t = \text{Tokens in clones}$$

>>>>

When refactoring two fragments that differ by a few literals, called methods or used variables, we are faced with a design tradeoff. For instance, if two literals differ and we want to refactor by method extraction, we need our new method to have an added argument to which each usage can pass the literal that was used at that place. Likewise, such things can be done for different variables that have been used and different called methods. Important here is, that the type of the used literals/variables are equal. Likewise, the signature of the called methods must be equal.

To limit the negative impact of this tradeoff on the design of the system, we formalized as threshold for the variability between fragments. The formula by which this threshold is calculated is displayed in equation \ref{eq:type2r}. We take the amount of literals, variables and method calls that differ from other clone instances in a clone class and divide it by the total number of tokens in the clone instance. Based on this threshold, we decide whether a clone should be considered for refactoring.

\begin{equation}\label{eq:type2r}
\text{T2R}=\frac{Diff(l) + Diff(v) + Diff(m)}{Total(t)}*100
\end{equation}
$$Diff = \text{Number that differ from other clone instances}$$
$$Total = \text{Total number in the clone instance}$$
$$l = \text{Literals in clone instance}$$
$$v = \text{Variables in clone instance}$$
$$m = \text{Method calls in clone instance}$$
$$t = \text{Tokens in clone instance}$$

A method call does often not consist of a single token (unlike variables and literals). For instance, a method call \texttt{System.out.println()} consists of several segments: a type reference to the \texttt{System} type, a reference to the static \texttt{out} field and a call of the \texttt{println()} method. Because of this, T2R clones are not a subset of of type 2 clones. Because T3R clones inherit all rules of T2R clones, T3R clones are also not a subset of T3 clones. This is displayed in equation \ref{eq:typerelation2}.
